Some examples in extended logic counting classes

### 2017-01-12 LICS/appendix-syntactic.tex
12 en

To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
$$
\sem{\alpha}(\A) = 
\begin{cases}
1 &\A \text{ has one element}\\
0 &\text{ otherwise}.
\end{cases}
$$
Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$. Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$. Since $\alpha$ takes the value 0 for some structures, $c$ must be 0. Consider a structure with one element $\mathfrak{1}$. We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$. Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}. Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.

@@

We separate this proof in three parts

\vspace{1em}
The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$. This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$. We base our proof on this fact.

For the first part, let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF. This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF. Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches. For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case. For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$. We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$. Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches. Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more. The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$. We conclude that $\alpha\in\totp$.

\vspace{1em}
For the second part, let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF. This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF. Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise. Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}. If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results. We conclude that $\alpha$ has an FPRAS.

\vspace{2em}
For the third part, we will prove that $\eqso(\loge{1})$ is closed under sum and multiplication. Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication. We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.

@@

\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication. This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.

\vspace{1em}
\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS. We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}. First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$. Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.

Let $\alpha$ be in $\eqso(\logex{0})$. The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$. Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$. Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$. We now define the function $g_{\alpha}$ procedurally. Let $\A$ be a $\R$-structure with domain $A$. Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty. For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$. If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$). This concludes the construction of $\A'$. Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure. Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$. We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$. Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.

To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}. Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}. These procedures also takes polynomial time and satisfies the condition.

\vspace{1em}
\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$. 
Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$. Moreover, we assume that $\length(\bar{x}) > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.

The proof will be separated in two parts. In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$. Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$. 
From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula. In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF. 
We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$. From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.

@@

\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ where $\varphi$ is quantifier-free and with  atomic formulas in $\fo$. We call this syntax {\em extended quantifier-free} from now on. 
Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ (i.e. $\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition. 
Therefore, for the rest of the proof we can assume that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.

 Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas. 
Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
$$
\varphi_i(\bar{X},\bar{x},\bar{y}) = \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
$$
Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.

Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) each first-order variable in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ is mentioned at most once and, moreover, no variable in $\bar{x}$ is mentioned, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
\begin{itemize}
	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$. We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$. We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$. The formula $\varphi_i$ is now defined as
	$$
	\varphi_i(\bar{X},\bar{x},(\bar{y},\bar{z})) = (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y}).
	$$
	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$. We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
	$$
	\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
	$$
	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$ 
	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$. For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$. A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$. On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.

>>>>

To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
$$
\sem{\alpha}(\A) = 
\begin{cases}
1 &\A \text{ has one element}\\
0 &\text{ otherwise}.
\end{cases}
$$
Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$. Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$. Since $\alpha$ takes the value 0 for some structures, $c$ must be 0. Consider a structure $\mathfrak{1}$ with one element. We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$. Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}. Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.

@@

The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$. This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$. We base our proof on this fact.

\vspace{1em}
$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF. This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF. Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches. For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case. For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$. We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$. Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches. Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more. The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$. We conclude that $\alpha\in\totp$.

\vspace{1em}
{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF. This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF. Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise. Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}. If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results. We conclude that $\alpha$ has an FPRAS.

\vspace{1em}
$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication. We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.

@@

{\bf Part 1.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication. This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.

\vspace{1em}
{\bf Part 2.} We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS. We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}. First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$. Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.

Let $\alpha$ be in $\eqso(\logex{0})$. The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$. Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$. Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$. We now define the function $g_{\alpha}$ procedurally. Let $\A$ be a $\R$-structure with domain $A$. Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty. For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$. If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$. This concludes the construction of $\A'$. Note that the number of $\fo$ subformulas and each arity and tuple size is fixed to $\alpha$, so computing this function takes polynomial time over the size of the structure. Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$. We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$. Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.

To show that $\alpha$ is in $\totp$, consider a procedure that takes an input $\enc(\A)$, converts it to $\enc(g_{\alpha}(\A))$ and simulates the $\totp$ procedure for $\lambda(\alpha)$ on input $\enc(g_{\alpha}(\A))$. This procedure generates exactly $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$ branches, and therefore $\alpha$ is in $\totp$.

To show that $\alpha$ has an FPRAS, consider a procedure that takes an input $\enc(\A)$, converts it to $\enc(g_{\alpha}(\A))$ and simulates the FPRAS for $\lambda(\alpha)$. This procedure also takes polynomial time and satisfies the condition.

\vspace{1em}
{\bf Part 3.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$. We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$. We suppose that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$. Moreover, we assume that $\length(\bar{x}) > 0$ and $\length(\bar{X}) > 0$ since we can replace each $\varphi_i$ for the equivalent formula $\sa{Y}\sa{y}\varphi_i\wedge\first(y)\wedge\forall z Y(z)$.

The proof will be separated in two parts. (i) If $\alpha$ is in $\logex{1}$-PNF, this is, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$, then we define a formula $\varphi'$ that satisfies the following condition. For each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$. We define $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$. (ii) If $\alpha = \beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$, then we define a formula $\varphi'$ that satisfies the following condition. For each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$ {\bf and} $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$. We define $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$.

@@

{\em Section (i).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ where $\varphi$ is quantifier-free with the addition of atomic formulas in $\fo$. We call this syntax {\em extended quantifier-free} from now on. Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
We convert $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m\varphi_i$ where each $\varphi_i$ is of the form:
$$
\varphi_i(\bar{X},\bar{x},\bar{y}) = \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge
\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}} \wedge \underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}}.
$$
This can be done with a standard predicate-DNF transformation algorithm. We now define a series of transformations to $\varphi$ that will make each formula $\varphi$ satisfy the following three conditions. (a) Each first-order variable in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ is mentioned at most once. Moreover, in this formula no variable in $\bar{x}$ is mentioned. (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$. We will define then what we mean by an ordered partition. (c) If $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
\begin{itemize}
	\item[(a)] Consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$. We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$. We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$. The formula $\varphi_i$ is now defined as
	$$
	\varphi_i(\bar{X},\bar{x},(\bar{y},\bar{z})) = (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y}).
	$$
	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$. We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
	$$
	\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
	$$
	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$ 
	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$. For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$. A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$. On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
